<!-- R Commander Markdown Template -->

Replace with Main Title
=======================

### BOUKA

### `r as.character(Sys.Date())`

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, out.width=750, fig.height=8, fig.width=8)
library(Rcmdr)
library(car)
library(RcmdrMisc)
```

```{r}
Dataset <- read.table("C:/Users/BOUKA/Downloads/Datos_Radiacion_1.txt", header=TRUE, stringsAsFactors=TRUE, 
  sep="\t", na.strings="NA", dec=".", strip.white=TRUE)
```

```{r}
Dataset <- read.table("C:/Users/BOUKA/Downloads/Datos_Radiacion_1.txt", header=TRUE, stringsAsFactors=TRUE, 
  sep="\t", na.strings="NA", dec=",", strip.white=TRUE)
```


### Linear Model: RegModel.2: Abateria~DJ+HR+NHL+RD+RDF+RG+Tm+Vv
```{r}
RegModel.2 <- lm(Abateria~DJ+HR+NHL+RD+RDF+RG+Tm+Vv, data=Dataset)
summary(RegModel.2)
```

```{r}
oldpar <- par(oma=c(0,0,3,0), mfrow=c(2,2))
```


```{r}
plot(RegModel.2)
```



### Basic Model Plots: RegModel.2
```{r}
par(oldpar)
```

```{r}
Dataset <- read.table("C:/Users/BOUKA/Downloads/Datos_Radiacion_1.txt", header=TRUE, stringsAsFactors=TRUE, 
  sep="\t", na.strings="NA", dec=",", strip.white=TRUE)
```


```{r}
powerTransform(Dataset[,"Abateria"],start=NULL)
```

```{r}
nuevodato <-bcPower(Dataset[,"Abateria"], 0.6513367)
```


```{r}
nuevodato <-bcPower(Dataset[,"Abateria"], 0.6513367)
```


```{r}
nuevodato
```



### Correlation Test: DJ, RD
```{r}
with(Dataset, cor.test(DJ, RD, alternative="two.sided", method="pearson"))
```

```{r}
library(abind, pos=16)
```


```{r}
library(e1071, pos=17)
```



### Numerical Summaries: Dataset
```{r}
numSummary(Dataset[,"Abateria", drop=FALSE], statistics=c("mean", "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,
  .75,1))
```


### Normality Test: ~Abateria
```{r}
normalityTest(~Abateria, test="shapiro.test", data=Dataset)
```


### Linear Model: RegModel.3: Abateria~NHL+Tm
```{r}
RegModel.3 <- lm(Abateria~NHL+Tm, data=Dataset)
summary(RegModel.3)
```


### Correlation Test: DJ, HR
```{r}
with(Dataset, cor.test(DJ, HR, alternative="two.sided", method="pearson"))
```

```{r}
model000 <- lm(Abateria ~ Tm + NHL - 1, data = Dataset)
```


```{r}
summary(model000)
```


```{r}
Dataset <- read.table("C:/Users/BOUKA/Downloads/Datos_Radiacion_2.txt", header=TRUE, stringsAsFactors=TRUE, 
  sep="\t", na.strings="NA", dec=",", strip.white=TRUE)
```


### Correlation Test: Abateria, RDF
```{r}
with(Dataset, cor.test(Abateria, RDF, alternative="two.sided", method="pearson"))
```


### Linear Model: RegModel.5: Abateria~NHL+RD+Tm
```{r}
RegModel.5 <- lm(Abateria~NHL+RD+Tm, data=Dataset)
summary(RegModel.5)
```

```{r}
RegModelR2 <- lm(Abateria~ NHL + RD + Tm -1, data=Dataset)
```


```{r}
summary(RegModelR2)
```



### Linear Model: RegModel.6: Abateria~DJ+HR+NHL+RD+RDF+RG+Tm+Vv
```{r}
RegModel.6 <- lm(Abateria~DJ+HR+NHL+RD+RDF+RG+Tm+Vv, data=Dataset)
summary(RegModel.6)
```


### Numerical Summaries: Dataset
```{r}
numSummary(Dataset[,"Tm", drop=FALSE], statistics=c("mean", "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```


### Normality Test: ~Abateria
```{r}
normalityTest(~Abateria, test="shapiro.test", data=Dataset)
```


### Numerical Summaries: Dataset
```{r}
numSummary(Dataset[,"Abateria", drop=FALSE], statistics=c("mean", "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,
  .75,1))
```


### Numerical Summaries: Dataset
```{r}
numSummary(Dataset[,"RG", drop=FALSE], statistics=c("mean", "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```


### Numerical Summaries: Dataset
```{r}
numSummary(Dataset[,"RDF", drop=FALSE], statistics=c("mean", "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```


### Linear Model: RegModel.7: Abateria~NHL
```{r}
RegModel.7 <- lm(Abateria~NHL, data=Dataset)
summary(RegModel.7)
```


### Correlation Test: Abateria, HR
```{r}
with(Dataset, cor.test(Abateria, HR, alternative="two.sided", method="pearson"))
```


### Correlation Test: HR, NHL
```{r}
with(Dataset, cor.test(HR, NHL, alternative="two.sided", method="pearson"))
```


### Correlation Test: DJ, RD
```{r}
with(Dataset, cor.test(DJ, RD, alternative="two.sided", method="pearson"))
```


### Correlation Test: RD, Tm
```{r}
with(Dataset, cor.test(RD, Tm, alternative="two.sided", method="pearson"))
```


### Correlation Test: RDF, RG
```{r}
with(Dataset, cor.test(RDF, RG, alternative="two.sided", method="pearson"))
```

```{r}
powerTransform(Dataset[,"Abateria"],start=NULL)
```


